---
title: "Twitter followers"
author: "Emily Darling"
date: "December 25, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=8, fig.path='Figs/',
                      echo=FALSE, warning=FALSE, message=FALSE)
library(dplyr)
library(reshape2)
library(stringr)
library(ggplot2)
library(ggrepel)
library(RColorBrewer)
library(vegan)
library(readxl)
library(grid)
library(gtable)
library(scales)
library(gridExtra)
library(extrafont)

```

## Twitter followers analysis

```{r read data, results = "hide"}
d <- read.csv(file.path(PROJHOME,"sci-twitter","data",
                   "output - all.data_7Dec2016.csv"),
         header = TRUE, stringsAsFactors = FALSE, strip.white = TRUE)
names(d)
d$handle <- tolower(d$handle)

scis110 <- read.csv(file.path(PROJHOME,"sci-twitter","data",
                              "110 handles_with Twitter dates.csv"),
                    header = TRUE, strip.white = TRUE, stringsAsFactors = FALSE)
names(scis110)[2] <- "handle"
#puttolower to match data analysis
scis110$handle <- tolower(scis110$handle)

subset(scis110, Followers > 5000)

head(scis110)
hist(scis110$Months.total)
hist(scis110$Year.joined)
```

This analysis looks at how academics accumulate Twitter followers, and types of followers, during their social media 'career'. This dataset comprises the `r nrow(d)` unique followers of `r length(unique(d$handle))` full-time academics. 

Followers have been classified into several major 'types' of followers using coded keywords and combinations of keywords. The accuracy of classification was checked by hand, and resulted in ~89% correct classification (~11% misclassification rate). 

## Preliminary plots

NUVI provided us with a reach of each follower (along with their bio). Here, for each academic Twitter handle, I've summed their total reach across each academic's followers. Not surprisingly, reach increases exponentially with the number of followers (note the log10 x-axis). Maybe we could do a GAM or breakpoint analysis to look if there's a threshold here for when your total reach starts to increase? 

```{r frequency of followers, results="hide", fig.width=7, fig.height=4.5}
head(d)
d2 <- d %>% 
  group_by(handle) %>% 
  summarize(median.reach = median(Reach, na.rm = TRUE),
            sum.reach = sum(Reach, na.rm = TRUE),
            max.reach = max(Reach, na.rm = TRUE))

levels(as.factor(d2$handle))

#check big reach numbers, yep, makes sense with 110 followers_long.xlsx
#and summarized in new SOM table
#filter(d, Reach > 4000000)

scis110 <- left_join(scis110, d2)
head(scis110)
names(scis110)
reach.melt <- melt(scis110[,c(2,4,13,14:16)], id.vars = 1:3)
head(reach.melt)

#plot of followers x total reach of Twitter academic
reach.melt$variable <- factor(reach.melt$variable, 
                              levels = c("median.reach",
                                         "max.reach",
                                         "sum.reach"))

reach.melt$variable <- recode_factor(reach.melt$variable, 
                              median.reach = "Median reach",
                              max.reach = "Maximum reach",
                              sum.reach = "Cumulative reach")

facet_names <- c(`Median reach` = "Median reach (average quality)",
                 `Maximum reach` = "Maximum reach (highest quality)",
                 `Cumulative reach` = "Cumulative reach (sum of quality)")

#plot - followers by reach
ggplot(aes(x = Followers, y = value), data = reach.melt) + 
  geom_point(size = 3, alpha = 0.75) + 
  theme_bw(base_size = 14) + 
  scale_x_log10(breaks = c(0,10,25,50,100,250,500,750,1000,
                           1500,2500,3500,5000,7500,10000)) +
  scale_y_continuous(labels = comma) +
  ylab("Twitter reach (follower's followers)") + 
  xlab("No. of followers") +
  stat_smooth(colour = "black", se = TRUE, fill = "grey50") + 
  theme(axis.text.x = element_text(angle = 45, hjust=1, vjust=1),
        text = element_text(size = 12),
        panel.grid = element_blank()) +
  facet_wrap(~variable, ncol = 1, scales = "free_y", 
             labeller = as_labeller(facet_names))

ggsave(file.path(PROJHOME,"paper","outputs",
                 "followers x reach.pdf"),
       height = 8, width = 5.5)

#histogram of followers
head(scis110)

ggplot(data = scis110) + 
  geom_histogram(aes(x = Followers), colour = "black", fill = "grey60",
                 boundary = 0) + 
  scale_x_log10(breaks = c(0,10,25,50,100,250,500,750,1000,
                           1500,2500,3500,5000,7500,10000)) +
  scale_y_continuous("Frequency", limits = c(0,13), expand = c(0,0),
                     breaks = c(0,5,10)) +
  theme_bw(base_size = 20) +
  theme(axis.text.x = element_text(angle = 45, hjust=1, vjust=1),
        axis.ticks.x = element_blank(),
        text = element_text(size = 12),
        panel.grid = element_blank(),
        plot.background = element_rect(fill = "transparent")) + 
  annotate("text", label = "n = 110 academics", x = 3000, y = 12)

ggsave(file.path(PROJHOME,"paper","outputs",
                 "followers log-histogram.pdf"),
       height = 3.15, width = 5.12)

#histogram of reach of followers
head(d)
d.reach <- d %>% 
  group_by(Username) %>% 
  summarize(Reach = mean(Reach))
min(d.reach$Reach); max(d.reach$Reach)
median(d.reach$Reach); sd(d.reach$Reach)

ggplot(data = d.reach) + 
  geom_histogram(aes(x = Reach), colour = "black", fill = "grey60",
                 boundary = 0) + 
  scale_x_log10(labels = comma,
                breaks = c(10,100,1000,10000,100000,1000000)) +
  scale_y_continuous("Frequency", limits = c(0,9000), expand = c(0,0)) + 
  theme_bw(base_size = 14) +
  theme(panel.grid = element_blank(),
        plot.background = element_rect(fill = "transparent")) + 
  annotate("text", label = "n = 64,666 followers", x = 900000, y = 7500)

ggsave(file.path(PROJHOME,"paper","outputs",
                 "reach log-histogram.pdf"),
       height = 3.15, width = 10)

#consider time and followers / reach
head(scis110)
head(reach.melt)

ggplot(aes(x = Months.total, y = value), data = reach.melt) + 
  geom_point(size = 3, alpha = 0.75) + 
  theme_bw(base_size = 14) +
  scale_y_continuous(labels = comma) +
  ylab("Twitter reach (follower's followers)") + 
  xlab("Months on Twitter") +
  stat_smooth(colour = "black", se = TRUE, fill = "grey50") + 
  theme(axis.text.x = element_text(angle = 45, hjust=1, vjust=1),
        text = element_text(size = 12),
        panel.grid = element_blank()) +
  facet_wrap(~variable, ncol = 1, scales = "free_y", 
             labeller = as_labeller(facet_names))

plot(scis110$Followers ~ scis110$Months.total,
     xlab = "Months total on Twitter",
     ylab = "Total followers")
abline(lm(scis110$Followers ~ scis110$Months.total))
summary(lm(scis110$Followers ~ scis110$Months.total))
text(x = 60, y = 8000, labels = "R2 = 0.04\nP= 0.02")

hist(scis110$Months.total)
```

Next, I've calculated the average reach for the followers classified to each group. 

```{r reach x group, results = "hide",fig.width=7, fig.height=4.5}
#1. summarize groups for each handle, from d

d2 <- d %>% 
  select(c(2,4:5,20,7:17)) %>% 
  melt(id.vars = c(1:4), variable.name = "group") %>% 
  filter(value > 0)

head(d2)
head(d2[which(is.na(d2$group)),])

#64,666 unique followers
length(unique(d2$Username))

#remove foreign publics, unlikely to be properly classified
d2 <- d2[-which(d2$foreign == 1 & d2$group == "public"),]
length(unique(d2$Username))
62958/64666

table(d2$group)

#62,958 unique followers once foreign public (unclassified) profiles are removed
length(unique(d2$Username)) 

#plot reach histograms for each group
head(d2)

#recode outreach (educators + mza)
d2$group <- recode_factor(d2$group, 
                          mza = "outreach")
levels(as.factor(d2$group))

d2$group <- factor(d2$group, levels = c("faculty", "student",
                                        "other.sci", "sci.assoc",
                                        "outreach","applied", 
                                        "media", "public","politician",
                                        "unknown"))

#combine scientists group
d2$group2 <- recode_factor(d2$group,
                                  faculty = "scientists",
                                  student = "scientists",
                                  other.sci = "scientists",
                                  sci.assoc = "scientists")
levels(as.factor(d2$group2))

#recode for outputs
d2$group <- recode_factor(d2$group, 
                          faculty = "Science faculty",
                          student = "Science students",
                          'other.sci' = "Other scientists",
                          'sci.assoc' = "Science associations",
                          outreach = "Outreach",
                          applied = "Applied",
                          media = "Media",
                          public = "Public",
                          politician = "Decision makers",
                          unknown = "Unknown")
levels(as.factor(d2$group))

d2$group2 <- recode_factor(d2$group2, 
                          scientists = "Scientists",
                          outreach = "Outreach",
                          applied = "Applied",
                          media = "Media",
                          public = "Public",
                          politician = "Decision makers",
                          unknown = "Unknown")
levels(as.factor(d2$group2))

## ==================================================
#boxplots by group
## ==================================================
head(d2)
length(unique(d2$Username))

d3 <- d2 %>% 
  group_by(handle) %>% 
  mutate(n_followers = n()) %>% 
  group_by(handle,group) %>% 
  summarize(n_group = n(),
            n_followers = mean(n_followers),
            prop_followers = n_group / n_followers,
            mean_reach = median(Reach),
            max_reach = max(Reach))
  
head(d3)
  
#boxplots by group - all scientists
# The palette with black:
levels(as.factor(d3$group))

group.colours <- c("#E69F00","#E69F00","#E69F00","#E69F00",
                   "yellow", "#009E73",
                   "dodgerblue","deepskyblue",
                   "navyblue","#999999")

group.colours.orig <- c("#0072B2","#0072B2","#0072B2","#0072B2",
                   "#56B4E9", "#009E73","#CC79A7",
                   "#D55E00","#E69F00",
                   "#999999")

unique(d3$group)
head(d3)
group.a <- ggplot(data = filter(d3,group != "Unknown"),
                  aes(x = reorder(group, -mean_reach), y = mean_reach)) +
  geom_boxplot(aes(colour = group)) + 
  scale_colour_manual(values = group.colours) +
  scale_y_log10(labels = comma,
                breaks = c(10,100,1000,10000,100000,1000000)) +
  theme_bw(base_size = 12) + 
  theme(panel.grid = element_blank(),
        legend.position = "none",
        axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        text=element_text(family="Times")) + 
    ylab("Median Twitter reach")
group.a

#mean reach by group, averaged across 110 handles
group.b <- ggplot(data = filter(d3,group != "Unknown"), 
                  aes(x = reorder(group, -mean_reach), y = prop_followers)) +
  geom_boxplot(aes(colour = group)) + 
  scale_colour_manual(values = group.colours) +
  scale_y_continuous(limits = c(0,0.45)) +
  theme_bw(base_size = 12) +
  theme(panel.grid = element_blank(),
        plot.background = element_rect(fill = "transparent"),
        axis.text.x = element_text(angle = 45, hjust=1, vjust=1),
        axis.title.x = element_blank(),
        legend.position = "none",
        text=element_text(family="Times")) + 
  ylab("Mean proportion of followers")
group.b

g <- arrangeGrob(rbind(ggplotGrob(group.a), ggplotGrob(group.b), size = "first"))
ggsave(file.path(PROJHOME,"paper","outputs",
                 "FigSX - boxplot proportion, reach x group.pdf"), g,
       height = 6, width = 6)


## ==================================================
#boxplots by group -  scientists grouped
## ==================================================
head(d2)
d4 <- d2 %>% 
  group_by(handle) %>% 
  mutate(n_followers = n()) %>% 
  group_by(handle,group2) %>% 
  summarize(n_group = n(),
            n_followers = mean(n_followers),
            prop_followers = n_group / n_followers,
            mean_reach = median(Reach))
  
head(d4)
  
levels(as.factor(d4$group2))
group2.colours <- c("#E69F00", "yellow", "#009E73",
                    "dodgerblue","deepskyblue","navyblue",
                    "#999999")

group2.colours.orig <- c("#0072B2", "#56B4E9", "#009E73",
                    "#CC79A7","#D55E00","#E69F00",
                   "#999999")

#create dummy facet of prop followers and reach 
head(d4)

group2.a <- ggplot(data = filter(d4, group2 != "Unknown"), 
                   aes(x = reorder(group2, -mean_reach), y = mean_reach)) +
  geom_boxplot(aes(colour = group2)) + 
  scale_colour_manual(values = group2.colours) + 
  scale_y_log10(labels = comma,
                breaks = c(10,100,1000,10000,100000,1000000)) +
  theme_bw(base_size = 20) + 
  theme(panel.grid = element_blank(),
        legend.position = "none",
        axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        text=element_text(family="Times")) + 
    ylab("Median Twitter reach")

#mean reach by group, averaged across 110 handles
group2.b <- ggplot(data = filter(d4, group2 != "Unknown"), 
                   aes(x = reorder(group2, -mean_reach), y = prop_followers)) +
  geom_boxplot(aes(colour = group2)) + 
  scale_colour_manual(values = group2.colours) +
  scale_y_continuous(limits = c(0,0.8)) + 
  theme_bw(base_size = 20) +
  theme(panel.grid = element_blank(),
        plot.background = element_rect(fill = "transparent"),
        axis.text.x = element_text(size = 18, angle = 45, hjust=1, vjust=1),
        axis.title.x = element_blank(),
        legend.position = "none",
        text=element_text(family="Times")) + 
  ylab("Mean proportion of followers") 


g <- arrangeGrob(rbind(ggplotGrob(group2.a), ggplotGrob(group2.b), size = "first"))
ggsave(file.path(PROJHOME,"paper","outputs",
                 "Fig3 - boxplot proportion, reach x group.pdf"), g,
       height = 8, width = 4.5)

  
#histograms of reach by group
unique(d2$group)
ggplot(data = dplyr::filter(d2, group != "Unknown")) + 
  geom_histogram(aes(x = Reach, fill = group), colour = "black",
                 boundary = 0) + 
  scale_x_log10(labels = comma,
                breaks = c(10,100,1000,10000,100000,1000000)) +
  theme_bw(base_size = 14) +
  theme(panel.grid = element_blank(),
        plot.background = element_rect(fill = "transparent"),
        axis.text.x = element_text(angle = 45, hjust=1, vjust=1),
        legend.position = "none") + 
  facet_wrap(~group, ncol = 3, scales = "free_y") + 
  scale_fill_manual(values = group.colours) +
  ylab("Frequency") + 
  xlab("Twitter reach")

ggsave(file.path(PROJHOME,"paper","outputs",
                 "reach x group log-histogram.pdf"),
       height = 6, width = 8)

#ANOVA of reach by groups
head(d4)
unique(d4$group2)

data <- dplyr::filter(d4, group2 != "Unknown")
model <- lm(data$mean_reach ~ data$group2)
anova(model)
kruskal.test(data$mean_reach ~ data$group2)
TukeyHSD(aov(data$mean_reach ~ data$group2))

model <- lm(data$prop_followers ~ data$group2)
anova(model)
kruskal.test(data$prop_followers ~ data$group2)
TukeyHSD(aov(data$prop_followers ~ data$group2))
```

We can then look at how the number of followers predicts which groups you accumulate (if we consider # of followers as a space-for-time comparison).


```{r followers x group, results = "hide",fig.width=9, fig.height=4.5}
head(d2)
length(unique(d2$Username))

## ==================================================
# follower accumulation plots -  scientists subgroups
## ==================================================
d5 <- d2 %>% 
  group_by(handle,group) %>% 
  summarize(n_group = n(),
            mean_reach = median(Reach),
            max_reach = max(Reach))
  
d5 <- left_join(scis110[,c(2:4)], d5)
head(d5)

levels(as.factor(d5$group))

#all levels on one plot
ggplot(data = filter(d5, group != "Unknown"), 
       aes(x = Followers, y = n_group)) + 
  geom_point(shape = 21, aes(colour = group, size = max_reach), 
             fill = "white", alpha = 0.5) + 
  stat_smooth(aes(colour = group), se = FALSE) + 
  theme_bw(base_size = 12) + 
  scale_x_log10(breaks = c(10,100,500,1000,2500,5000,10000)) + 
  scale_colour_manual("Group", values = group.colours) +
  theme(panel.grid = element_blank(),
        text=element_text(family="Times")) + 
  ylab("No. of group followers") +
  xlab("log10 Twitter followers") +
  scale_size_continuous("Maximum \nreach", labels = comma)

ggsave(file.path(PROJHOME,"paper","outputs",
                 "follower increase x group.pdf"),
       height = 5, width = 8)

#multi plot of follower increase by group
ggplot(data = filter(d5, group != "Unknown"), 
       aes(x = Followers, y = n_group)) + 
  stat_smooth(aes(colour = group), se = FALSE) + 
  geom_point(shape = 21, aes(colour = group, size = max_reach), 
             fill = "white", alpha = 0.5) + 
  theme_bw(base_size = 12) + 
  scale_x_log10(breaks = c(100,1000,10000)) + 
  scale_colour_manual("Group", values = group.colours) +
  facet_wrap(~ group) + 
  theme(panel.grid = element_blank(),
        text=element_text(family="Times")) + 
  ylab("Group followers") +
  xlab("Twitter followers") +
  scale_size_continuous("Maximum \nreach", labels = comma)

ggsave(file.path(PROJHOME,"paper","outputs",
                 "multiplot - follower increase x group.pdf"),
       height = 6, width = 8.5)

## ==================================================
# follower accumulation plots -  scientists combined (group2)
## ==================================================
head(d2)
d6 <- d2 %>% 
  group_by(handle,group2) %>% 
  summarize(n_group2 = n(),
            mean_reach = median(Reach),
            max_reach = max(Reach))
  
d6 <- left_join(scis110[,c(2:4)], d6)
head(d6)

levels(as.factor(d6$group2))

#all levels on one plot
head(d6)
ggplot(data = filter(d6, group2 != "Unknown"), 
       aes(x = Followers, y = n_group2)) + 
  geom_point(shape = 21, aes(colour = group2, size = max_reach), 
             fill = "white", alpha = 0.5) + 
  stat_smooth(aes(colour = group2), se = FALSE) + 
  theme_bw(base_size = 20) + 
  scale_x_log10(breaks = c(10,100,500,1000,2500,5000,10000)) + 
  scale_colour_manual("Group", values = group2.colours) +
  theme(panel.grid = element_blank(),
        text=element_text(family="Times")) + 
  ylab("Group followers") +
  xlab("Total followers") +
  scale_size_continuous("Maximum \nreach", labels = comma)

ggsave(file.path(PROJHOME,"paper","outputs",
                 "Fig2 - follower increase x group2.pdf"),
       height = 5, width = 10)

#multi plot of follower increase by group2
ggplot(data = filter(d6, group2 != "Unknown"), 
       aes(x = Followers, y = n_group2)) + 
  stat_smooth(aes(colour = group2), se = FALSE) + 
  geom_point(shape = 21, aes(colour = group2, size = max_reach), 
             fill = "white", alpha = 0.5) + 
  theme_bw(base_size = 14) + 
  scale_x_log10(breaks = c(100,1000,10000)) + 
  scale_colour_manual("Group", values = group2.colours) +
  facet_wrap(~ group2, scales = "free_y") + 
  theme(panel.grid = element_blank(),
        text=element_text(family="Times"),
        legend.position = "none",
        axis.title = element_blank()) + 
  ylab("Group followers") +
  xlab("Total followers") +
  scale_size_continuous("Maximum \nreach", labels = comma)

ggsave(file.path(PROJHOME,"paper","outputs",
                 "multiplot - follower increase x group2.pdf"),
       height = 4, width = 7)

## does maximum reach (e.g., high quality followers) increase with followers? 

## MODELS
#fit reach x group
#public and media significantly accrue higher quality reaches with more followers
library(nlme)
head(d6)
m0 <- gls(max_reach ~ Followers * group2,
          data = d6,
          na.action = na.exclude)
summary(m0)

ggplot(data = filter(d6, group2 != "Unknown"), 
       aes(x = Followers, y = max_reach)) + 
  geom_point(shape = 21, aes(colour = group2, size = max_reach), 
             fill = "white", alpha = 0.5) + 
  stat_smooth(aes(colour = group2), se = FALSE, method = "lm") + 
  theme_bw(base_size = 10) + 
  scale_x_log10(breaks = c(10,100,500,1000,2500,5000,10000)) + 
  scale_y_continuous(labels = comma) +
  scale_colour_manual("Group", values = group2.colours) +
  theme(panel.grid = element_blank()) + 
  ylab("Maximum reach") +
  xlab("log10 Twitter followers") +
  scale_size_continuous("Maximum \nreach", labels = comma)


ggsave(file.path(PROJHOME,"paper","outputs",
                 "model - reach x followers increase.pdf"),
       height = 5, width = 6.5)
```


